---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---
```{r intro, echo=FALSE, results="hide"}
##### MUST CHANGE MANUALLY TWO LINES: 38 & 42 (date of data) #####
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 
#knitr::opts_knit$set(root.dir = "C:/Users/yoonjoung.choi/Dropbox/0 iSquared/iSquared_Harfod/1.COVID19/")
knitr::opts_knit$set(root.dir = "C:/Users/yoonjoung choi/Dropbox/0 iSquared/iSquared_Harfod/1.COVID19/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(data.table)))

suppressWarnings(suppressMessages(library(readxl)))
suppressWarnings(suppressMessages(library(Matrix))) 
suppressWarnings(suppressMessages(library(stringr)))
suppressWarnings(suppressMessages(library(lubridate)))
suppressWarnings(suppressMessages(library(zoo)))

suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(plotly)))
suppressWarnings(suppressMessages(library(knitr)))  
```

# COVID-19 vaccine coverage in Harford County, MD

(Page updated at `r time`)   
Data as of: 03/23/2021, from Maryland Department of Health. See Annex for detailed note about the data.    

```{r dta_import}
#Immunization level data 
#dtaimmunet<-read_excel("C:/Users/yoonjoung.choi/Documents/dta_immunet/COVID-19 Administrations - Harford 03232021.xlsx")
dtaimmunet<-read_excel("C:/Users/yoonjoung choi/Dropbox/0 iSquared/iSquared_Harfod/dta_immunet/COVID-19 Administrations - Harford 03232021.xlsx")


dim(dtaimmunet)
length(unique(dtaimmunet$CLIENT_ID))
```

```{r dta_checkclean}
# Check duplicates, total number of vaccination, etc.  
dta<-as.data.frame(dtaimmunet) %>% 
    mutate(vdate	=as.Date(VACCINATION_DATE, "%d-%b-%Y"))%>%
    arrange(CLIENT_ID, vdate)%>%
    group_by(CLIENT_ID)%>% 
    mutate(
        dup=CLIENT_ID==lag(CLIENT_ID)  &
            vdate==lag(vdate), #duplicate
        totalnumber=n(), #Total number of vaccinations received per person
        dosenumber=row_number() #Vaccine dose number per person, should be 1-2 
        )%>%
    ungroup()

    table(dta$dup)    
    table(dta$totalnumber)
    table(dta$dosenumber)
    
#check people who got more than 2 doses, why??? 
    temp<-dta%>%
        filter(totalnumber>2)%>% 
        select(CLIENT_ID, VACCINATION_DATE, BIRTH_DATE, SEX_CODE, MANUFACTURER)
    head(temp, 10)
    
#check vaccine date AND replace incorrect date with missing
    summary(dta$vdate)
    temp<-dta%>%
        filter(vdate < as.Date("2020-12-01"))%>% 
        select(CLIENT_ID, VACCINATION_DATE, BIRTH_DATE, SEX_CODE, MANUFACTURER,
               DATE_ENTERED1, LAST_UPDATED_DATE1)
    head(temp, 10)    
    
nincorrectvdate<-nrow(temp)

dta<-dta%>%
    mutate(
        vdate = replace(vdate, vdate < as.Date("2020-12-01"), NA)
    ) 
    
```    

```{r dta_processing}
dta<-dta%>%
    arrange(CLIENT_ID, VACCINATION_DATE)%>%
    mutate(
                
        bday	=as.Date(BIRTH_DATE, "%d-%b-%Y"), 
        bdaybase   =as.Date("14-JAN-2021", "%d-%b-%Y"),  
        today   = as.Date(Sys.time(	), format='%d%b%Y'), 
        age     =floor((bdaybase-bday)/365), #AGE as of January 14 in complete years
        
        complete = (totalnumber>=2 & (MANUFACTURER=="MOD" | MANUFACTURER=="PFR") | 
                    totalnumber==1 & MANUFACTURER=="JSN"), #completed required doses
        waiting1 = complete==0 & 
                (MANUFACTURER=="MOD" | MANUFACTURER=="PFR") & 
                today-vdate<28, #waiting for MOD/PFR 2nd dose (<=28)
        waiting2 = complete==0 & 
                (MANUFACTURER=="MOD" | MANUFACTURER=="PFR") & 
                (today-vdate>=28 & today-vdate<=41), #waiting for MOD/PFR 2nd dose (29-42)   
        late = complete==0 & 
                (MANUFACTURER=="MOD" | MANUFACTURER=="PFR") & 
                today-vdate>=42, #late for MOD/PFR 2nd dose
            
        complete=ifelse((is.na(vdate)==TRUE | MANUFACTURER==""), NA, complete),
        waiting1=ifelse((is.na(vdate)==TRUE | MANUFACTURER==""), NA, waiting1),
        waiting2=ifelse((is.na(vdate)==TRUE | MANUFACTURER==""), NA, waiting2),
        late    =ifelse((is.na(vdate)==TRUE | MANUFACTURER==""), NA, late),

	    nhwhite = ETHNICITY_CODE=="2186-5" & RACE_CODE=="2106-3", #Non-Hispanic White"
	    nhblack = ETHNICITY_CODE=="2186-5" & RACE_CODE=="2054-5", #Non-Hispanic Black
	    nhasian = ETHNICITY_CODE=="2186-5" & RACE_CODE=="2028-9", #Non-Hispanic Asian
	    hispanic = ETHNICITY_CODE=="2135-2", #Hispanic 
	
    	zipcode17 = 
    		ZIP_CODE=="21001" |
    		ZIP_CODE=="21005" |
    		ZIP_CODE=="21009" |
    		ZIP_CODE=="21014" |
    		ZIP_CODE=="21015" |
    		ZIP_CODE=="21017" |
    		ZIP_CODE=="21028" |
    		ZIP_CODE=="21034" |
    		ZIP_CODE=="21040" |
    		ZIP_CODE=="21047" |
    		ZIP_CODE=="21050" |
    		ZIP_CODE=="21078" |
    		ZIP_CODE=="21084" |
    		ZIP_CODE=="21085" |
    		ZIP_CODE=="21132" |
    		ZIP_CODE=="21154" |
    		ZIP_CODE=="21160"  #"17 non-border zipcode areas"
        
    ) 
```

##1. Coverage: overall and by background characteristics  
###1.1 One or more doses
```{r coverage_CUM}
dtafirst<-dta%>%filter(dosenumber==1)
    
dta1<-dtafirst%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Total")

dta2<-dtafirst%>%filter(nhwhite==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic White")
dta3<-dtafirst%>%filter(nhblack==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Black")
dta4<-dtafirst%>%filter(nhasian==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Asian")
dta5<-dtafirst%>%filter(hispanic==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic")

dta6<-dtafirst%>%filter(SEX_CODE=="F")%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Female")
dta7<-dtafirst%>%filter(SEX_CODE=="M")%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Male")

dta8<-dtafirst%>%filter(age>=75)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="75+")
dta9<-dtafirst%>%filter(age>=65 & age<75)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="65-74")

dta10<-dtafirst%>%filter(age>=65 & nhwhite==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="White 65+")
dta11<-dtafirst%>%filter(age>=65 & nhblack==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Black 65+")
#dta12<-dtafirst%>%filter(age>=65 & nhasian==1)%>%
#    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
#    mutate(group="Asian 65+")
dta12<-dtafirst%>%filter(age>=65 & hispanic==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic 65+")

dtasummary<-rbind(dta1, dta2, dta3, dta4, dta5, 
                  dta6, dta7, dta8, dta9, dta10, dta11, dta12)%>%
    rename(count=dosenumber)

```

```{r coverage_7day}
dtafirst7day<-dta%>%filter(dosenumber==1)%>%filter(today - vdate<7)
    
dta1<-dtafirst7day%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Total")

dta2<-dtafirst7day%>%filter(nhwhite==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic White")
dta3<-dtafirst7day%>%filter(nhblack==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Black")
dta4<-dtafirst7day%>%filter(nhasian==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Asian")
dta5<-dtafirst7day%>%filter(hispanic==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic")

dta6<-dtafirst7day%>%filter(SEX_CODE=="F")%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Female")
dta7<-dtafirst7day%>%filter(SEX_CODE=="M")%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Male")

dta8<-dtafirst7day%>%filter(age>=75)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="75+")
dta9<-dtafirst7day%>%filter(age>=65 & age<75)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="65-74")

dta10<-dtafirst7day%>%filter(age>=65 & nhwhite==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="White 65+")
dta11<-dtafirst7day%>%filter(age>=65 & nhblack==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Black 65+")
#dta12<-dtafirst7day%>%filter(age>=65 & nhasian==1)%>%
#    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
#    mutate(group="Asian 65+")
dta12<-dtafirst7day%>%filter(age>=65 & hispanic==1)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic 65+")

dtasummary7day<-rbind(dta1, dta2, dta3, dta4, dta5, 
                  dta6, dta7, dta8, dta9, dta10, dta11, dta12)%>%
    rename(count7day=dosenumber)
```

```{r coverage_pop}
dtasummary<-left_join(dtasummary, dtasummary7day, by="group")%>%
    mutate(
        pop=NA, 
        pop = replace(pop, group=="Total", 	255441	),
        pop = replace(pop, group=="Non-Hispanic White", 	191517	),
        pop = replace(pop, group=="Non-Hispanic Black", 	35826	),
        pop = replace(pop, group=="Non-Hispanic Asian", 	7299	),
        pop = replace(pop, group=="Hispanic", 	12215	),
        pop = replace(pop, group=="Female", 	130676	),
        pop = replace(pop, group=="Male", 	124765	),
        pop = replace(pop, group=="75+", 	17417	),
        pop = replace(pop, group=="65-74", 	24869	),
        pop = replace(pop, group=="White 65+", 	36064	),
        pop = replace(pop, group=="Black 65+", 	5666	),
        pop = replace(pop, group=="Hispanic 65+", 	555	),

        coverage = round(100*count/pop), 
        coverage7day = round(100*count7day/pop, 1)

    )
```

```{r table_coverage, results="asis"}
setcolorder(dtasummary, c("group", "count", "coverage", "count7day", "coverage7day"))

kable(dtasummary[, 1:5], caption = "COVID-19 vaccine coverage, one or more doses (number of people and percent of population): cumulative and past 7 days")
```

###1.2 By status
```{r status}
dtafirst<-dta%>%filter(dosenumber==1) #person-level data
    
dta1<-dtafirst%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Total")

dta2<-dtafirst%>%filter(nhwhite==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic White")
dta3<-dtafirst%>%filter(nhblack==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Black")
dta4<-dtafirst%>%filter(nhasian==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Non-Hispanic Asian")
dta5<-dtafirst%>%filter(hispanic==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic")

dta6<-dtafirst%>%filter(SEX_CODE=="F")%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Female")
dta7<-dtafirst%>%filter(SEX_CODE=="M")%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Male")

dta8<-dtafirst%>%filter(age>=75)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="75+")
dta9<-dtafirst%>%filter(age>=65 & age<75)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="65-74")

dta10<-dtafirst%>%filter(age>=65 & nhwhite==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="White 65+")
dta11<-dtafirst%>%filter(age>=65 & nhblack==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Black 65+")
#dta12<-dtafirst%>%filter(age>=65 & nhasian==1)%>%
#    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
#    mutate(group="Asian 65+")
dta12<-dtafirst%>%filter(age>=65 & hispanic==1)%>%
    summarize_at(vars(dosenumber, complete, waiting1, waiting2, late), funs(sum(., na.rm = TRUE)))%>%
    mutate(group="Hispanic 65+")

dtasummarystatus<-rbind(dta1, dta2, dta3, dta4, dta5, 
                  dta6, dta7, dta8, dta9, dta10, dta11, dta12)%>%
    rename(total=dosenumber)

```

```{r status_pop}
dtasummarystatus<-dtasummarystatus%>%
    mutate(
        pop=NA, 
        pop = replace(pop, group=="Total", 	255441	),
        pop = replace(pop, group=="Non-Hispanic White", 	191517	),
        pop = replace(pop, group=="Non-Hispanic Black", 	35826	),
        pop = replace(pop, group=="Non-Hispanic Asian", 	7299	),
        pop = replace(pop, group=="Hispanic", 	12215	),
        pop = replace(pop, group=="Female", 	130676	),
        pop = replace(pop, group=="Male", 	124765	),
        pop = replace(pop, group=="75+", 	17417	),
        pop = replace(pop, group=="65-74", 	24869	),
        pop = replace(pop, group=="White 65+", 	36064	),
        pop = replace(pop, group=="Black 65+", 	5666	),
        pop = replace(pop, group=="Hispanic 65+", 	555	),

        #coverage = round(100*count/pop), 
        #coverage7day = round(100*count7day/pop, 1)
        pcttotal = round(100*total/pop), 
        pctcomplete = round(100*complete/pop), 
        pctwaiting1 = round(100*waiting1/pop), 
        pctwaiting2 = round(100*waiting2/pop), 
        pctlate = round(100*late/pop, 1)
    )
```

```{r table_status_count, results="asis"}
setcolorder(dtasummarystatus, c("group", "total", "complete", "waiting1", "waiting2", "late"))

kable(dtasummarystatus[, 1:6], caption = "COVID-19 vaccine coverage by status (number of people): complete, waiting adequate (<28 days), waiting late (28-41 days), vs. late")
```
Note:  
1. Complete: either two doses of Moderna or Pfizer, or one dose of J&J vaccine  
2. Sum of the number of complete, waiting1, waiting2, and late may not be same with the total number of people, due to a small number of cases with missing information on manufacture and/or vaccination date.   

```{r table_status_pct, results="asis"}
setcolorder(dtasummarystatus, c("group", "pcttotal", "pctcomplete", "pctwaiting1", "pctwaiting2", "pctlate"))

kable(dtasummarystatus[, 1:6], caption = "COVID-19 vaccine coverage by status (percent of population): complete, waiting adequate (<28 days), waiting late (28-41 days), vs. late")
```
Note:  
1. Complete: either two doses of Moderna or Pfizer, or one dose of J&J vaccine


```{r defaulters} 
dtadefaulter<-dta%>%
    filter(waiting2==1 | late==1)%>% #keep late or almost late folks 
    filter(COUNTY_OF_RESIDENCE=="Harford")%>% #Keep only HC residents
    mutate(
        status= " " ,
        status= replace(status, late==1, "late"), 
        status= replace(status, waiting2==1, "waiting late"), 
    
        race= " ", 
        race= replace(race, nhwhite==1, "NH White"), 
        race= replace(race, nhblack==1, "NH Black"), 
        race= replace(race, nhasian==1, "NH Asian"), 
        race= replace(race, hispanic==1, "Hispanic"), 
        
        days_since_first = today - vdate
        )%>%
    arrange(status, days_since_first)
    
setcolorder(dtadefaulter, c("status",  "days_since_first", "race", "age"))
colnames(dtadefaulter)

dtadefaulter<-dtadefaulter[, 1:33]

write.csv(dtadefaulter, paste0("HCHD_defaulters_", date, ".csv"))

```

##2. Coverage by zipcode among Harford county residents  
###2.1 One or more doses 
```{r coverage_Zip}
dtasummaryZip<-dta%>%
    filter(dosenumber==1)%>%
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    filter(zipcode17==1)%>%
    group_by(ZIP_CODE)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    rename(count=dosenumber)

dtasummaryZip7day<-dta%>%
    filter(dosenumber==1)%>%filter(today - vdate<7)%>%
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    filter(zipcode17==1)%>%
    group_by(ZIP_CODE)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    rename(count7day=dosenumber)

dtasummaryZip<-merge(dtasummaryZip, dtasummaryZip7day, by = "ZIP_CODE" )%>%
    mutate(
	    pop = NA, #"population, by zipcode"
        pop = replace(pop, ZIP_CODE=="21001"	 , 	24752	),
        pop = replace(pop, ZIP_CODE=="21005"	 , 	2872	),
        pop = replace(pop, ZIP_CODE=="21009"	 , 	29905	),
        pop = replace(pop, ZIP_CODE=="21013"	 , 	4768	),
        pop = replace(pop, ZIP_CODE=="21014"	 , 	36538	),
        pop = replace(pop, ZIP_CODE=="21015"	 , 	29199	),
        pop = replace(pop, ZIP_CODE=="21017"	 , 	6983	),
        pop = replace(pop, ZIP_CODE=="21028"	 , 	2783	),
        pop = replace(pop, ZIP_CODE=="21034"	 , 	3483	),
        pop = replace(pop, ZIP_CODE=="21040"	 , 	24166	),
        pop = replace(pop, ZIP_CODE=="21047"	 , 	12266	),
        pop = replace(pop, ZIP_CODE=="21050"	 , 	18123	),
        pop = replace(pop, ZIP_CODE=="21078"	 , 	18366	),
        pop = replace(pop, ZIP_CODE=="21082"	 , 	656	),
        pop = replace(pop, ZIP_CODE=="21084"	 , 	7603	),
        pop = replace(pop, ZIP_CODE=="21085"	 , 	16055	),
        pop = replace(pop, ZIP_CODE=="21087"	 , 	5394	),
        pop = replace(pop, ZIP_CODE=="21111"	 , 	5387	),
        pop = replace(pop, ZIP_CODE=="21132"	 , 	3232	),
        pop = replace(pop, ZIP_CODE=="21154"	 , 	5804	),
        pop = replace(pop, ZIP_CODE=="21161"	 , 	5601	),
        pop = replace(pop, ZIP_CODE=="21160"	 , 	2933	),
        
        coverage = round(100*count/pop), 
        coverage7day = round(100*count7day/pop, 1)
    )

setcolorder(dtasummaryZip, c("ZIP_CODE", "count", "coverage", "count7day", "coverage7day"))

```

```{r coveragedtaformap}
#####1. Get coverage data ready
# Full list of zipcode in Harford
HCzipcode <- read_excel("C:/Users/yoonjoung choi/Dropbox/0 iSquared/iSquared_Harfod/dta_demobasic/HCzipcode.xlsx")
names(HCzipcode)<-c("area", "ZCTA5CE10", "border") 

# Coverage data calculated above
dtamap<- dtasummaryZip%>%
    select(ZIP_CODE, coverage)%>%
    rename(ZCTA5CE10=ZIP_CODE)%>%
    mutate(ZCTA5CE10=as.numeric(ZCTA5CE10))

# Full list of zipcode merged with coverage data 
dtamap<-left_join(HCzipcode, dtamap, by = "ZCTA5CE10")

#####2. tigris: Download a shapefile (shp,gpkg,geojson...)
library(sf) #Overall handling of sf objects
library(tigris) #For downloading the zipcode map
options(tigris_use_cache = TRUE)
geo <- st_as_sf(zctas(cb = TRUE, starts_with = dta$ZCTA5CE10))

#Overall shape of USA states
states <- st_as_sf(states(cb=TRUE))

#For plotting, all the maps should have the same crs
states=st_transform(states,st_crs(geo))

#####3. Now Merge MY coverage data
dta.sf=merge(geo,dtamap)
```

```{r plot_coveragemap, results='asis'}
#4. Plotting
library(cartography) #Plotting maps package

par(mar=c(1,1,1,1))
#ghostLayer(dta.sf)

#plot(st_geometry(states), add=TRUE)

choroLayer(dta.sf,
           var="coverage",
           add=FALSE,
           border = "black",
           method = "quantile", nclass = 8,
           col = carto.pal(pal1="red.pal", n1=2,
                           pal2="blue.pal", n2=6),
           legend.title.txt = "Percent",
           legend.title.cex = 0.8, 
           legend.values.cex = 0.8,
           legend.pos = "bottomleft",
           legend.frame = FALSE)

layoutLayer(title = "COVID-19 vaccine coverage by zip code (one ore more doses): Harford County",
            col="white",
            coltitle = "black",
            scale = FALSE,
            frame = FALSE
            )
```

```{r table_coverageZip, results="asis"}
kable(dtasummaryZip[, 1:5], caption = "COVID-19 vaccine coverage, one or more doses by zip code (number of people and percent of population): cumulative and past 7 days")
```

###2.2 Complete doses
```{r coverage2_Zip}
dtasummaryZip<-dta%>%
    filter(complete==1)%>%
    mutate(dosenumber=1)%>% #replace dosenumber to 1 in all rows
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    filter(zipcode17==1)%>%
    group_by(ZIP_CODE)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    rename(count=dosenumber)

dtasummaryZip7day<-dta%>%
    filter(complete==1)%>%
    mutate(dosenumber=1)%>% #replace dosenumber to 1 in all rows
    filter(today - vdate<7)%>%
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    filter(zipcode17==1)%>%
    group_by(ZIP_CODE)%>%
    summarize_at(vars(dosenumber), funs(sum(., na.rm = TRUE)))%>%
    rename(count7day=dosenumber)

dtasummaryZip<-merge(dtasummaryZip, dtasummaryZip7day, by = "ZIP_CODE" )%>%
    mutate(
	    pop = NA, #"population, by zipcode"
        pop = replace(pop, ZIP_CODE=="21001"	 , 	24752	),
        pop = replace(pop, ZIP_CODE=="21005"	 , 	2872	),
        pop = replace(pop, ZIP_CODE=="21009"	 , 	29905	),
        pop = replace(pop, ZIP_CODE=="21013"	 , 	4768	),
        pop = replace(pop, ZIP_CODE=="21014"	 , 	36538	),
        pop = replace(pop, ZIP_CODE=="21015"	 , 	29199	),
        pop = replace(pop, ZIP_CODE=="21017"	 , 	6983	),
        pop = replace(pop, ZIP_CODE=="21028"	 , 	2783	),
        pop = replace(pop, ZIP_CODE=="21034"	 , 	3483	),
        pop = replace(pop, ZIP_CODE=="21040"	 , 	24166	),
        pop = replace(pop, ZIP_CODE=="21047"	 , 	12266	),
        pop = replace(pop, ZIP_CODE=="21050"	 , 	18123	),
        pop = replace(pop, ZIP_CODE=="21078"	 , 	18366	),
        pop = replace(pop, ZIP_CODE=="21082"	 , 	656	),
        pop = replace(pop, ZIP_CODE=="21084"	 , 	7603	),
        pop = replace(pop, ZIP_CODE=="21085"	 , 	16055	),
        pop = replace(pop, ZIP_CODE=="21087"	 , 	5394	),
        pop = replace(pop, ZIP_CODE=="21111"	 , 	5387	),
        pop = replace(pop, ZIP_CODE=="21132"	 , 	3232	),
        pop = replace(pop, ZIP_CODE=="21154"	 , 	5804	),
        pop = replace(pop, ZIP_CODE=="21161"	 , 	5601	),
        pop = replace(pop, ZIP_CODE=="21160"	 , 	2933	),
        
        coverage = round(100*count/pop), 
        coverage7day = round(100*count7day/pop, 1)
    )

setcolorder(dtasummaryZip, c("ZIP_CODE", "count", "coverage", "count7day", "coverage7day"))

```

```{r coverage2dtaformap}
#####1. Get coverage data ready
# Full list of zipcode in Harford
HCzipcode <- read_excel("C:/Users/yoonjoung choi/Dropbox/0 iSquared/iSquared_Harfod/dta_demobasic/HCzipcode.xlsx")
names(HCzipcode)<-c("area", "ZCTA5CE10", "border") 

# Coverage data calculated above
dtamap<- dtasummaryZip%>%
    select(ZIP_CODE, coverage)%>%
    rename(ZCTA5CE10=ZIP_CODE)%>%
    mutate(ZCTA5CE10=as.numeric(ZCTA5CE10))

# Full list of zipcode merged with coverage data 
dtamap<-left_join(HCzipcode, dtamap, by = "ZCTA5CE10")

#####2. tigris: Download a shapefile (shp,gpkg,geojson...)
library(sf) #Overall handling of sf objects
library(tigris) #For downloading the zipcode map
options(tigris_use_cache = TRUE)
geo <- st_as_sf(zctas(cb = TRUE, starts_with = dta$ZCTA5CE10))

#Overall shape of USA states
states <- st_as_sf(states(cb=TRUE))

#For plotting, all the maps should have the same crs
states=st_transform(states,st_crs(geo))

#####3. Now Merge MY coverage data
dta.sf=merge(geo,dtamap)
```

```{r plot_coverage2map, results='asis'}
#4. Plotting
library(cartography) #Plotting maps package

par(mar=c(1,1,1,1))
#ghostLayer(dta.sf)

#plot(st_geometry(states), add=TRUE)

choroLayer(dta.sf,
           var="coverage",
           add=FALSE,
           border = "black",
           method = "quantile", nclass = 8,
           col = carto.pal(pal1="red.pal", n1=2,
                           pal2="blue.pal", n2=6),
           legend.title.txt = "Percent",
           legend.title.cex = 0.8, 
           legend.values.cex = 0.8,
           legend.pos = "bottomleft",
           legend.frame = FALSE)

layoutLayer(title = "COVID-19 vaccine coverage by zip code (complete): Harford County",
            col="white",
            coltitle = "black",
            scale = FALSE,
            frame = FALSE
            )


```
Note:  
1. Complete: either two doses of Moderna or Pfizer, or one dose of J&J vaccine

```{r table_coverage2Zip, results='asis'}
kable(dtasummaryZip[, 1:5], caption = "COVID-19 vaccine coverage, complete by zip code (number of people and percent of population): cumulative and past 7 days")
```
Note:  
1. Complete: either two doses of Moderna or Pfizer, or one dose of J&J vaccine

##Annex: Detailed description of the data
```{r}
nobs<-nrow(dta)
npersons<-length(unique(dta$CLIENT_ID))
vdatefirst<-min(as.Date.POSIXct(dta$vdate, "1970-01-01"))
vdatelast<-max(as.Date.POSIXct(dta$vdate, "1970-01-01"))

nhcresidents<-dta%>%filter(dosenumber==1)%>%
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    nrow()

nzipcode17<-dta%>%filter(dosenumber==1)%>%
    filter(COUNTY_OF_RESIDENCE=="Harford")%>%
    filter(zipcode17==1)%>%
    nrow()
```
As of `r date`  
- A total of `r nobs` vaccination cases from `r npersons` people.        
- A total of `r nhcresidents` Harford county residents and `r nzipcode17` living in the 17 zip code areas are included in the data.   
- Vaccination date ranges from `r vdatefirst` to `r vdatelast`.       

- Number of vaccination cases with missing entry or likely errors (i.e., implausible values)    
* Missing race/ethnicity:    
* Missing age:   
* Missing sex:   
* Likely error in vaccine date: `r nincorrectvdate`   

---

<p style="color:gray">
See [GitHub](https://github.com/yoonjoung/COVID19_FlattenedCurve) for data, code, and more information. 
For typos, errors, and questions, contact me at [www.isquared.global](https://www.iSquared.global/YJ). 

_Making Data Delicious, One Byte at a Time_</p>